---
title: "Validation of Extremes"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Code setup

Loading libraries and functions, reading and preprocessing data.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(heatwaveR)
library(ncdf4)
library(R.matlab)
library(patchwork)
library(logKDE)
library(philentropy)
```

```{r}
read_data = FALSE
process = "sst"
sites <- c("Port l'Hebert", "Taylor's Head Shallow")
site_labels <- c("Port L'Hebert", "Taylor's Head")
pal <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442")
source("setup.R")
```

# Plotting FVCOM time series dates

```{r}
hourly <- hourly_dataframe |>
  filter(between(time, dates$extreme[1], dates$extreme[2])) |>
  select(-GLORYS) |>
  rename("t" = time) |>
  pivot_longer(
    c(OBS, FVCOM, FVCOM2),
    names_to = "type",
    values_to = "temp"
  ) |>
  remove_missing_times()

hourly |>
  ggplot(aes(x = t, y = temp, col = type)) +
  geom_line() +
  facet_wrap(~site, ncol = 1)

daily <- daily_dataframe |>
  filter(between(time, dates$extreme[1], dates$extreme[2])) |>
  select(-GLORYS) |>
  rename("t" = time) |>
  pivot_longer(
    c(OBS, FVCOM, FVCOM2),
    names_to = "type",
    values_to = "temp"
  ) |>
  remove_missing_times()

daily |>
  ggplot(aes(x = t, y = temp, col = type)) +
  geom_line() +
  facet_wrap(~site, ncol = 1)
```

# Port L'Hebert Traditional MHWs

Note that using hourly data we do not get the same MHWs. For the analysis, we should use the daily data to match the Hobday definition as close as possible.

## Daily OBS

```{r}
ph_obs_daily_clm <- daily |>
  filter(site == "Port L'Hebert") |>
  filter(type == "OBS") |>
  select(-type) |>
  mhw_climatology() |>
  detect_events()

ph_obs_daily_events <- ph_obs_daily_clm |>
  calculate_events(funcs = event_funcs)

ph_obs_daily_clm |>
  pivot_longer(
    c(temp, seas, thresh),
    names_to = "temp_type",
    values_to = "temp"
  ) |>
  ggplot(aes(x = t, y = temp, col = temp_type)) +
  geom_line()

ph_obs_daily_events
```

## Daily FVCOM

```{r}
ph_fvcom_daily_clm <- daily |>
  filter(site == "Port L'Hebert") |>
  filter(type == "FVCOM") |>
  select(-type) |>
  mhw_climatology() |>
  detect_events()

ph_fvcom_daily_events <- ph_fvcom_daily_clm |>
  calculate_events(funcs = event_funcs)

ph_fvcom_daily_clm |>
  pivot_longer(
    c(temp, seas, thresh),
    names_to = "temp_type",
    values_to = "temp"
  ) |>
  ggplot(aes(x = t, y = temp, col = temp_type)) +
  geom_line()

ph_fvcom_daily_events

# 2018-07-29 NOT IN FVCOM2
# 2020-11-12 SHOWN
# 2020-12-02
# 2021-09-26
# 2021-10-18 AVAILABLE
```

## Daily FVCOMv2

```{r}
ph_fvcom2_daily_clm <- daily |>
  filter(site == "Port L'Hebert") |>
  filter(type == "FVCOM2") |>
  select(-type) |>
  mhw_climatology() |>
  detect_events()

ph_fvcom2_daily_events <- ph_fvcom2_daily_clm |>
  calculate_events(funcs = event_funcs)

ph_fvcom2_daily_clm |>
  pivot_longer(
    c(temp, seas, thresh),
    names_to = "temp_type",
    values_to = "temp"
  ) |>
  ggplot(aes(x = t, y = temp, col = temp_type)) +
  geom_line()

ph_fvcom2_daily_events

# 2020-09-29 NOT IN FVCOM
# 2020-11-12 SHOWN
# 2021-03-25
# 2021-10-15 AVAILABLE
```

## Comparing Events (2020-11-12)

```{r}
tibble(
  "data" = c("Observations", "FVCOMv1", "FVCOMv2"),
  "events" = c(nrow(ph_obs_daily_events),
               nrow(ph_fvcom_daily_events),
               nrow(ph_fvcom2_daily_events)),
  "mean_duration" = c(mean(ph_obs_daily_events$duration),
                      mean(ph_fvcom_daily_events$duration),
                      mean(ph_fvcom2_daily_events$duration)),
  "mean_intensity" = c(mean(ph_obs_daily_events$intensity_mean),
                       mean(ph_fvcom_daily_events$intensity_mean),
                       mean(ph_fvcom2_daily_events$intensity_mean)),
  "mean_cumulative_intensity" = c(mean(ph_obs_daily_events$intensity_cumulative),
                                  mean(ph_fvcom_daily_events$intensity_cumulative),
                                  mean(ph_fvcom2_daily_events$intensity_cumulative))
)

#compare_all_events(ph_obs_daily_clm, ph_fvcom_daily_clm, ymd("2020-10-15"),
#                   ymd("2020-12-15"), label1 = "Observations", label2 = "FVCOMv1")

compare_event(
  clms = list(ph_obs_daily_clm, ph_fvcom_daily_clm, ph_fvcom2_daily_clm),
  nums = c(3, 5, 3),
  labels = c("Observations", "FVCOMv1", "FVCOMv2"),
  start_date = ymd("2020-10-23"),
  end_date = ymd("2020-12-02")
) +
  scale_x_datetime(date_labels = "%Y-%m") +
  theme(
    legend.position = "right",
    legend.background = element_rect(color = "white"),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(lwd = 1.2))) +
  labs(
    title = "",
    x = "Time"
  )
```

```{r}
compare_all_events(
  clms = list(ph_obs_daily_clm, ph_fvcom_daily_clm, ph_fvcom2_daily_clm),
  labels = c("Observations", "FVCOMv1", "FVCOMv2"),
  start_date = ymd("2020-01-01"),
  end_date = ymd("2021-12-31")
) +
  scale_x_datetime(date_labels = "%Y-%m") +
  theme(
    legend.position = "right",
    legend.background = element_rect(color = "white"),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(lwd = 1.2))) +
  labs(
    title = "",
    x = "Time"
  )
```

```{r}
compare_exceedances(
  clms = list(ph_obs_daily_clm, ph_fvcom_daily_clm, ph_fvcom2_daily_clm),
  labels = c("Observations", "FVCOMv1", "FVCOMv2"),
  start_date = ymd("2020-06-01"),
  end_date = ymd("2021-12-31")
) +
  scale_x_datetime(date_labels = "%Y-%m") +
  theme(
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  labs(
    title = "",
    x = "Time"
  )
```


# Taylor's Head Traditional MHWs

## Daily OBS

```{r}
th_obs_daily_clm <- daily |>
  filter(site == "Taylor's Head") |>
  filter(type == "OBS") |>
  select(-type) |>
  mhw_climatology() |>
  detect_events()

th_obs_daily_events <- th_obs_daily_clm |>
  calculate_events(funcs = event_funcs)

th_obs_daily_clm |>
  pivot_longer(
    c(temp, seas, thresh),
    names_to = "temp_type",
    values_to = "temp"
  ) |>
  ggplot(aes(x = t, y = temp, col = temp_type)) +
  geom_line()

th_obs_daily_events
```

## Daily FVCOM

```{r}
th_fvcom_daily_clm <- daily |>
  filter(site == "Taylor's Head") |>
  filter(type == "FVCOM") |>
  select(-type) |>
  mhw_climatology() |>
  detect_events()

th_fvcom_daily_events <- th_fvcom_daily_clm |>
  calculate_events(funcs = event_funcs)

th_fvcom_daily_clm |>
  pivot_longer(
    c(temp, seas, thresh),
    names_to = "temp_type",
    values_to = "temp"
  ) |>
  ggplot(aes(x = t, y = temp, col = temp_type)) +
  geom_line()

th_fvcom_daily_events
```

## Daily FVCOMv2

```{r}
th_fvcom2_daily_clm <- daily |>
  filter(site == "Taylor's Head") |>
  filter(type == "FVCOM2") |>
  select(-type) |>
  mhw_climatology() |>
  detect_events()

th_fvcom2_daily_events <- th_fvcom2_daily_clm |>
  calculate_events(funcs = event_funcs)

th_fvcom2_daily_clm |>
  pivot_longer(
    c(temp, seas, thresh),
    names_to = "temp_type",
    values_to = "temp"
  ) |>
  ggplot(aes(x = t, y = temp, col = temp_type)) +
  geom_line()

th_fvcom2_daily_events
```

## Comparing Events (2021-02-06)

```{r}
tibble(
  "data" = c("Observations", "FVCOMv1", "FVCOMv2"),
  "events" = c(nrow(th_obs_daily_events),
               nrow(th_fvcom_daily_events),
               nrow(th_fvcom2_daily_events)),
  "mean_duration" = c(mean(th_obs_daily_events$duration),
                      mean(th_fvcom_daily_events$duration),
                      mean(th_fvcom2_daily_events$duration)),
  "mean_intensity" = c(mean(th_obs_daily_events$intensity_mean),
                       mean(th_fvcom_daily_events$intensity_mean),
                       mean(th_fvcom2_daily_events$intensity_mean)),
  "mean_cumulative_intensity" = c(mean(th_obs_daily_events$intensity_cumulative),
                                  mean(th_fvcom_daily_events$intensity_cumulative),
                                  mean(th_fvcom2_daily_events$intensity_cumulative))
)

#compare_all_events(th_obs_daily_clm, th_fvcom_daily_clm, ymd("2020-10-15"),
#                   ymd("2020-12-15"), label1 = "Observations", label2 = "FVCOMv1")

compare_event(
  clms = list(th_obs_daily_clm, th_fvcom_daily_clm, th_fvcom2_daily_clm),
  nums = c(4, 4, 5),
  labels = c("Observations", "FVCOMv1", "FVCOMv2"),
  start_date = ymd("2020-11-16"),
  end_date = ymd("2020-12-26")
) +
  scale_x_datetime(date_labels = "%Y-%m") +
  theme(
    legend.position = "right",
    legend.background = element_rect(color = "white"),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(lwd = 1.2))) +
  labs(
    title = "",
    x = "Time"
  )

# 2020-12-06 4,4,5
# 2021-02-08 5,6,7
# 2021-10-23 8,9,10
```

```{r}
compare_exceedances(
  clms = list(th_obs_daily_clm, th_fvcom_daily_clm, th_fvcom2_daily_clm),
  labels = c("Observations", "FVCOMv1", "FVCOMv2"),
  start_date = ymd("2020-06-01"),
  end_date = ymd("2021-12-31")
) +
  scale_x_datetime(date_labels = "%Y-%m") +
  theme(
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  labs(
    title = "",
    x = "Time"
  )
```

# Port L'Hebert EVT

Start with using a spline climatology with 90% threshold. Then we will try to use parameter stability plots to determine appropriate thresholds. Start with a runtime of $r=3$ to see if that produces enough extremes

## Parameter stability plots

```{r}
runtime <- 3

ph_obs_clm <- hourly |>
  filter(site == "Port L'Hebert") |>
  filter(type == "OBS") |>
  select(-type) |>
  periodic_climatology2(
    filter_func = smooth.spline,
    filter_params = list(nknots = 11),
    extract = "y",
    hourly = TRUE,
    padding = 30*24,
    thresh_quantile = 0.9
  )

ph_obs_quantile_thresh <- ph_obs_clm$thresh[1] - ph_obs_clm$seas[1]
ph_obs_anomalies <- na.omit(ph_obs_clm$temp - ph_obs_clm$seas)
fit_thresholds(ph_obs_anomalies, range = c(1, 3.5), 
               nout = 30, alpha = 0.05) +
  labs(
    title = "",
    y = "Estimated parameters"
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

ph_fvcom_clm <- hourly |>
  filter(site == "Port L'Hebert") |>
  filter(type == "FVCOM") |>
  select(-type) |>
  periodic_climatology2(
    filter_func = smooth.spline,
    filter_params = list(nknots = 11),
    extract = "y",
    hourly = TRUE,
    padding = 30*24,
    thresh_quantile = 0.9
  )

ph_fvcom_quantile_thresh <- ph_fvcom_clm$thresh[1] - ph_fvcom_clm$seas[1]
ph_fvcom_anomalies <- na.omit(ph_fvcom_clm$temp - ph_fvcom_clm$seas)
fit_thresholds(ph_fvcom_anomalies, range = c(1, 2), 
               nout = 30, alpha = 0.05, init = c(1,0)) +
  labs(
    title = "",
    y = "Estimated parameters"
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

ph_fvcom2_clm <- hourly |>
  filter(site == "Port L'Hebert") |>
  filter(type == "FVCOM2") |>
  select(-type) |>
  periodic_climatology2(
    filter_func = smooth.spline,
    filter_params = list(nknots = 11),
    extract = "y",
    hourly = TRUE,
    padding = 30*24,
    thresh_quantile = 0.9
  )

ph_fvcom2_quantile_thresh <- ph_fvcom2_clm$thresh[1] - ph_fvcom2_clm$seas[1]
ph_fvcom2_anomalies <- na.omit(ph_fvcom2_clm$temp - ph_fvcom2_clm$seas)
fit_thresholds(ph_fvcom2_anomalies, range = c(1, 3.7), 
               nout = 30, alpha = 0.05, init = c(0,1)) +
  labs(
    title = "",
    y = "Estimated parameters"
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

# Threshold choices: 2, 1.4, 2
```

## Calculating exceedances

```{r}
ph_obs_clm <- ph_obs_clm |>
  mutate(thresh = seas + 2) |>
  detect_events(min_duration = 1, max_gap = runtime)
ph_obs_events <- calculate_events(ph_obs_clm, funcs = event_funcs, target_col = "exceedance")

ph_fvcom_clm <- ph_fvcom_clm |>
  mutate(thresh = seas + 1.5) |>
  detect_events(min_duration = 1, max_gap = runtime)
ph_fvcom_events <- calculate_events(ph_fvcom_clm, funcs = event_funcs, target_col = "exceedance")

ph_fvcom2_clm <- ph_fvcom2_clm |>
  mutate(thresh = seas + 2) |>
  detect_events(min_duration = 1, max_gap = runtime)
ph_fvcom2_events <- calculate_events(ph_fvcom2_clm, funcs = event_funcs, target_col = "exceedance")


ph_obs_exceedance_plot <- plot_exceedances(ph_obs_clm,
                                           start_date = ymd("2018-04-01"),
                                           end_date = ymd("2018-06-30")) +
  labs(
    title = "Observation Exceedances at Port L'Hebert"
  )
ph_obs_metric_plot <- plot_event_metric(ph_obs_events, ind = "date_peak", 
                                        metric = "exceedance_max",
                                        start_ind = ymd("2018-04-01"),
                                        end_ind = ymd("2018-06-30")) +
  labs(
    title = "Declustered Observation Exceedances at Port L'Hebert"
  )

ph_obs_events
ph_fvcom_events
ph_fvcom2_events
ph_obs_exceedance_plot / ph_obs_metric_plot
```

## Plotting distribution functions

```{r}
ph_obs_exceedances <- ph_obs_events$exceedance_max
ph_obs_mle <- gp_mle(ph_obs_exceedances, init = c(0,0))

ph_fvcom_exceedances <- ph_fvcom_events$exceedance_max
ph_fvcom_mle <- gp_mle(ph_fvcom_exceedances, init = c(0,0))

ph_fvcom2_exceedances <- ph_fvcom2_events$exceedance_max
ph_fvcom2_mle <- gp_mle(ph_fvcom2_exceedances, init = c(1,0))

plot_gp_hist(ph_obs_exceedances, ph_obs_mle, "black", 0.7) +
  labs(
    x = "Exceedance [\u00B0C]",
    title = ""
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

plot_gp_hist(ph_fvcom_exceedances, ph_fvcom_mle, pal[1], 0.7) +
  labs(
    x = "Exceedance [\u00B0C]",
    title = ""
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

plot_gp_hist(ph_fvcom2_exceedances, ph_fvcom2_mle, pal[2], 0.7) +
  labs(
    x = "Exceedance [\u00B0C]",
    title = ""
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )
```

## Plotting all distribution functions

```{r}
n_ph_obs <- length(ph_obs_exceedances)
n_ph_fvcom <- length(ph_fvcom_exceedances)
n_ph_fvcom2 <- length(ph_fvcom2_exceedances)
n_th_obs <- length(th_obs_exceedances)
n_th_fvcom <- length(th_fvcom_exceedances)
n_th_fvcom2 <- length(th_fvcom2_exceedances)

tibble(
  "exceedance" = c(
    ph_obs_exceedances, ph_fvcom_exceedances, ph_fvcom2_exceedances,
    th_obs_exceedances, th_fvcom_exceedances, th_fvcom2_exceedances
  ),
  "site" = c(
    rep("Port L'Hebert", n_ph_obs),
    rep("Port L'Hebert", n_ph_fvcom),
    rep("Port L'Hebert", n_ph_fvcom2),
    rep("Taylor's Head", n_th_obs),
    rep("Taylor's Head", n_th_fvcom),
    rep("Taylor's Head", n_th_fvcom2)
  ),
  "type" = c(
    rep("OBS", n_ph_obs),
    rep("FVCOM", n_ph_fvcom),
    rep("FVCOM2", n_ph_fvcom2),
    rep("OBS", n_th_obs),
    rep("FVCOM", n_th_fvcom),
    rep("FVCOM2", n_th_fvcom2)
  )
) |>
  group_by(site, type) |>
  mutate("max_exceedance" = max(exceedance)) |>
  ungroup() |>
  mutate("type" = factor(type, levels = c("OBS", "FVCOM", "FVCOM2"))) |>
  ggplot(aes(x = exceedance)) +
  geom_histogram(aes(y = after_stat(density)),
                 breaks = seq(0, max_exceedance, length.out = 30),
                 fill = "lightgrey", alpha = 0.5, color = "black") +
  facet_grid(type ~ site, 
             scales = "free",
             labeller = as_labeller(c("OBS" = "Observations",
                                      "FVCOM" = "FVCOMv1",
                                      "FVCOM2" = "FVCOMv2",
                                      "Port L'Hebert" = "Port L'Hebert",
                                      "Taylor's Head" = "Taylor's Head"))) +
  theme_bw()


x_range <- seq(from = 0, to = max(x), length.out = 100)
f <- gp_density(mle$par, x_range)
  
hist_data <- tibble(x)
dist_data <- tibble("x" = x_range, "f" = f)
  
hist_data |>
  ggplot(aes(x)) +
  geom_histogram(aes(y = after_stat(density)),
                 breaks = seq(0, max(x), length.out = bins),
                 fill = "lightgrey", alpha = 0.5, color = "black") +  
  geom_line(data = dist_data, aes(x = x, y = f), color = "blue") +
  labs(
    title = "Generalized Pareto Modeled Distribution Function",
    y = "Density"
  ) +
  theme_bw()
```
## Comparing parameter estimates

```{r}
params <- c(ph_obs_mle$par, ph_fvcom_mle$par, ph_fvcom2_mle$par)
se <- c(ph_obs_mle$se, ph_fvcom_mle$se, ph_fvcom2_mle$se)

tibble(
  "value" = params,
  "upper" = params + se,
  "lower" = params - se,
  "param" = rep(c("Scale", "Shape"), 3),
  "type" = factor(rep(c("Observations", "FVCOMv1", "FVCOMv2"), each = 2), 
                  levels = c("Observations", "FVCOMv1", "FVCOMv2"))
) |>
  ggplot(aes(x = type, y = value, col = type)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1,
                position = position_dodge(0.05)) +
  facet_wrap(~param, ncol = 2, scales = "free") +
  scale_color_manual(
    name = "",
      values = c("black", pal[1], pal[2]),
      breaks = c("Observations", "FVCOMv1", "FVCOMv2"),
      labels = c("Observations", "FVCOMv1", "FVCOMv2")
  ) +
  theme_bw() +
  labs(
    title = "",
    x = "Data",
    y = "Parameter value"
  ) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )
```

## Calculating KL Divergence

```{r}
x <- seq(0, 100, length.out = 10000)
ph_obs_density <- gp_density(ph_obs_mle$par, x)
ph_fvcom_density <- gp_density(ph_fvcom_mle$par, x)
ph_fvcom2_density <- gp_density(ph_fvcom2_mle$par, x)

ph_obs_density <- ph_obs_density / sum(ph_obs_density)
ph_fvcom_density <- ph_fvcom_density / sum(ph_fvcom_density)
ph_fvcom2_density <- ph_fvcom2_density / sum(ph_fvcom2_density)

KL(rbind(ph_obs_density, ph_fvcom_density), unit = "log")
KL(rbind(ph_obs_density, ph_fvcom2_density), unit = "log")
```

## Calculating "excess heat"

```{r}
exceedance_area(ph_obs_clm)
exceedance_area(ph_fvcom_clm)
exceedance_area(ph_fvcom2_clm)
```

## Comparing EV distributions

```{r}
tibble(
  "x" = seq(0, 6, length.out = 1000),
  "obs" = gp_density(ph_obs_mle$par, x),
  "fvcom" = gp_density(ph_fvcom_mle$par, x),
  "fvcom2" = gp_density(ph_fvcom2_mle$par, x)
) |>
  pivot_longer(
    c(obs, fvcom, fvcom2),
    names_to = "type",
    values_to = "density"
  ) |>
  ggplot(aes(x, y = density, col = type)) +
  geom_line() +
  scale_color_manual(
      name = "",
      values = c("black", pal[1], pal[2]),
      breaks = c("obs", "fvcom", "fvcom2"),
      labels = c("Observations", "FVCOMv1", "FVCOMv2")
    ) +
  labs(
    title = "",
    y = "Density",
    x = "Exceedance [\u00B0C]",
    color = ""
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    legend.key.width = unit(1.5, "cm"),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(lwd = 1.2)))
```


## Trying the same thing but with cumulative exceedance (GETTING NEGATIVE CUMULATIVE EXCEEDANCES FOR SOME REASON!!!)

```{r}
ph_obs_exceedance_plot <- plot_exceedances(ph_obs_clm,
                                           start_date = ymd("2018-04-01"),
                                           end_date = ymd("2018-06-30")) +
  labs(
    title = "Observation Exceedances at Port L'Hebert"
  )
ph_obs_metric_plot <- plot_event_metric(ph_obs_events, ind = "date_peak", 
                                        metric = "exceedance_cumulative",
                                        start_ind = ymd("2018-04-01"),
                                        end_ind = ymd("2018-06-30")) +
  labs(
    title = "Observation Declustered Exceedances at Port L'Hebert"
  )

ph_obs_events
ph_fvcom_events
ph_fvcom2_events
ph_obs_exceedance_plot / ph_obs_metric_plot
```

```{r}
ph_obs_exceedances_cumulative <- ph_obs_events$exceedance_cumulative
ph_obs_mle_cumulative <- gp_mle(ph_obs_exceedances_cumulative, init = c(0,0))

ph_fvcom_exceedances_cumulative <- ph_fvcom_events$exceedance_cumulative
ph_fvcom_mle_cumulative <- gp_mle(ph_fvcom_exceedances_cumulative, init = c(0,0))

ph_fvcom2_exceedances_cumulative <- ph_fvcom2_events$exceedance_cumulative
ph_fvcom2_mle_cumulative <- gp_mle(ph_fvcom2_exceedances_cumulative, init = c(0,0))

plot_gp_hist(ph_obs_exceedances_cumulative, ph_obs_mle_cumulative)
plot_gp_empirical(ph_obs_exceedances_cumulative, ph_obs_mle_cumulative)

plot_gp_hist(ph_fvcom_exceedances_cumulative, ph_fvcom_mle_cumulative)
plot_gp_hist(ph_fvcom2_exceedances_cumulative, ph_fvcom2_mle_cumulative)
```

```{r}
params_cumulative <- c(ph_obs_mle_cumulative$par, 
                       ph_fvcom_mle_cumulative$par, 
                       ph_fvcom2_mle_cumulative$par)
se_cumulative <- c(ph_obs_mle_cumulative$se, 
                   ph_fvcom_mle_cumulative$se, 
                   ph_fvcom2_mle_cumulative$se)

tibble(
  "value" = params_cumulative,
  "upper" = params_cumulative + se_cumulative,
  "lower" = params_cumulative - se_cumulative,
  "param" = rep(c("Scale", "Shape"), 3),
  "type" = factor(rep(c("Observations", "FVCOMv1", "FVCOMv2"), each = 2), 
                  levels = c("Observations", "FVCOMv1", "FVCOMv2"))
) |>
  ggplot(aes(x = type, y = value)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1,
                position = position_dodge(0.05)) +
  facet_wrap(~param, ncol = 1, scales = "free") +
  theme_bw()
```










# Taylor's Head EVT

## Parameter Stability Plots

```{r}
runtime <- 3

th_obs_clm <- hourly |>
  filter(site == "Taylor's Head") |>
  filter(type == "OBS") |>
  select(-type) |>
  periodic_climatology2(
    filter_func = smooth.spline,
    filter_params = list(nknots = 11),
    extract = "y",
    hourly = TRUE,
    padding = 30*24,
    thresh_quantile = 0.9
  )

th_obs_quantile_thresh <- th_obs_clm$thresh[1] - th_obs_clm$seas[1]
th_obs_anomalies <- na.omit(th_obs_clm$temp - th_obs_clm$seas)
fit_thresholds(th_obs_anomalies, range = c(1, 2.5), 
               nout = 30, alpha = 0.05) +
  labs(
    title = "",
    y = "Estimated parameters"
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

th_fvcom_clm <- hourly |>
  filter(site == "Taylor's Head") |>
  filter(type == "FVCOM") |>
  select(-type) |>
  periodic_climatology2(
    filter_func = smooth.spline,
    filter_params = list(nknots = 11),
    extract = "y",
    hourly = TRUE,
    padding = 30*24,
    thresh_quantile = 0.9
  )

th_fvcom_quantile_thresh <- th_fvcom_clm$thresh[1] - th_fvcom_clm$seas[1]
th_fvcom_anomalies <- na.omit(th_fvcom_clm$temp - th_fvcom_clm$seas)
fit_thresholds(th_fvcom_anomalies, range = c(1, 2.2), 
               nout = 30, alpha = 0.05, init = c(1,0)) +
  labs(
    title = "",
    y = "Estimated parameters"
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

th_fvcom2_clm <- hourly |>
  filter(site == "Taylor's Head") |>
  filter(type == "FVCOM2") |>
  select(-type) |>
  periodic_climatology2(
    filter_func = smooth.spline,
    filter_params = list(nknots = 11),
    extract = "y",
    hourly = TRUE,
    padding = 30*24,
    thresh_quantile = 0.9
  )

th_fvcom2_quantile_thresh <- th_fvcom2_clm$thresh[1] - th_fvcom2_clm$seas[1]
th_fvcom2_anomalies <- na.omit(th_fvcom2_clm$temp - th_fvcom2_clm$seas)
fit_thresholds(th_fvcom2_anomalies, range = c(1, 2.1), 
               nout = 30, alpha = 0.05, init = c(1,1)) +
  labs(
    title = "",
    y = "Estimated parameters"
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

# Threshold choices: 1.6, 1.5, 1.7
```

## Calculating exceedances

```{r}
th_obs_clm <- th_obs_clm |>
  mutate(thresh = seas + 1.7) |>
  detect_events(min_duration = 1, max_gap = runtime)
th_obs_events <- calculate_events(th_obs_clm, funcs = event_funcs, target_col = "exceedance")

th_fvcom_clm <- th_fvcom_clm |>
  mutate(thresh = seas + 1.5) |>
  detect_events(min_duration = 1, max_gap = runtime)
th_fvcom_events <- calculate_events(th_fvcom_clm, funcs = event_funcs, target_col = "exceedance")

th_fvcom2_clm <- th_fvcom2_clm |>
  mutate(thresh = seas + 1.7) |>
  detect_events(min_duration = 1, max_gap = runtime)
th_fvcom2_events <- calculate_events(th_fvcom2_clm, funcs = event_funcs, target_col = "exceedance")


th_obs_exceedance_plot <- plot_exceedances(th_obs_clm,
                                           start_date = ymd("2018-04-01"),
                                           end_date = ymd("2018-06-30")) +
  labs(
    title = "Observation Exceedances at Taylor's Head"
  )
th_obs_metric_plot <- plot_event_metric(th_obs_events, ind = "date_peak", 
                                        metric = "exceedance_max",
                                        start_ind = ymd("2018-04-01"),
                                        end_ind = ymd("2018-06-30")) +
  labs(
    title = "Observation Declustered Exceedances at Taylor's Head"
  )

th_obs_events
th_fvcom_events
th_fvcom2_events
th_obs_exceedance_plot / th_obs_metric_plot
```

## Plotting distribution functions

```{r}
th_obs_exceedances <- th_obs_events$exceedance_max
th_obs_mle <- gp_mle(th_obs_exceedances, init = c(0,0))

th_fvcom_exceedances <- th_fvcom_events$exceedance_max
th_fvcom_mle <- gp_mle(th_fvcom_exceedances, init = c(1,0))

th_fvcom2_exceedances <- th_fvcom2_events$exceedance_max
th_fvcom2_mle <- gp_mle(th_fvcom2_exceedances, init = c(0,0))

plot_gp_hist(th_obs_exceedances, th_obs_mle, "black", 0.7) +
  labs(
    x = "Exceedance [\u00B0C]",
    title = ""
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

plot_gp_hist(th_fvcom_exceedances, th_fvcom_mle, pal[1], 0.7) +
  labs(
    x = "Exceedance [\u00B0C]",
    title = ""
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

plot_gp_hist(th_fvcom2_exceedances, th_fvcom2_mle, pal[2], 0.7) +
  labs(
    x = "Exceedance [\u00B0C]",
    title = ""
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )
```

## Comparing parameter estimates

```{r}
params <- c(th_obs_mle$par, th_fvcom_mle$par, th_fvcom2_mle$par)
se <- c(th_obs_mle$se, th_fvcom_mle$se, th_fvcom2_mle$se)

tibble(
  "value" = params,
  "upper" = params + se,
  "lower" = params - se,
  "param" = rep(c("Scale", "Shape"), 3),
  "type" = factor(rep(c("Observations", "FVCOMv1", "FVCOMv2"), each = 2), 
                  levels = c("Observations", "FVCOMv1", "FVCOMv2"))
) |>
  ggplot(aes(x = type, y = value, color = type)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1,
                position = position_dodge(0.05)) +
  facet_wrap(~param, ncol = 2, scales = "free") +
  scale_color_manual(
    name = "",
      values = c("black", pal[1], pal[2]),
      breaks = c("Observations", "FVCOMv1", "FVCOMv2"),
      labels = c("Observations", "FVCOMv1", "FVCOMv2")
  ) +
  theme_bw() +
  labs(
    title = "",
    x = "Data",
    y = "Parameter value"
  ) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )
```

## Calculating KL divergence

```{r}
x <- seq(0, 10, length.out = 1000)
th_obs_density <- gp_density(th_obs_mle$par, x)
th_fvcom_density <- gp_density(th_fvcom_mle$par, x)
th_fvcom2_density <- gp_density(th_fvcom2_mle$par, x)

th_obs_density <- th_obs_density / sum(th_obs_density)
th_fvcom_density <- th_fvcom_density / sum(th_fvcom_density)
th_fvcom2_density <- th_fvcom2_density / sum(th_fvcom2_density)

KL(rbind(th_obs_density, th_fvcom_density), unit = "log")
KL(rbind(th_obs_density, th_fvcom2_density), unit = "log")
```

```{r}
exceedance_area(th_obs_clm)
exceedance_area(th_fvcom_clm)
exceedance_area(th_fvcom2_clm)
```

## Comparing EV distributions

```{r}
tibble(
  "x" = seq(0, 4.5, length.out = 1000),
  "obs" = gp_density(th_obs_mle$par, x),
  "fvcom" = gp_density(th_fvcom_mle$par, x),
  "fvcom2" = gp_density(th_fvcom2_mle$par, x)
) |>
  pivot_longer(
    c(obs, fvcom, fvcom2),
    names_to = "type",
    values_to = "density"
  ) |>
  ggplot(aes(x, y = density, col = type)) +
  geom_line() +
  scale_color_manual(
      name = "",
      values = c("black", pal[1], pal[2]),
      breaks = c("obs", "fvcom", "fvcom2"),
      labels = c("Observations", "FVCOMv1", "FVCOMv2")
    ) +
  labs(
    title = "",
    y = "Density",
    x = "Exceedance [\u00B0C]",
    color = ""
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    legend.key.width = unit(1.5, "cm"),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(lwd = 1.2)))
```




# Comparing GLORYS data

```{r}
ph_glorys <- daily_data$glorys |>
  filter(site == "Port L'Hebert") |>
  filter(between(t, dates$extreme[1], dates$extreme[2]))

th_glorys <- daily_data$glorys |>
  filter(site == "Taylor's Head") |>
  filter(between(t, dates$extreme[1], dates$extreme[2]))
```

```{r}
ph_glorys_clm <- ph_glorys |>
  mhw_climatology() |>
  detect_events()

ph_glorys_events <- ph_glorys_clm |>
  calculate_events(funcs = event_funcs)

ph_glorys_clm |>
  pivot_longer(
    c(temp, seas, thresh),
    names_to = "temp_type",
    values_to = "temp"
  ) |>
  ggplot(aes(x = t, y = temp, col = temp_type)) +
  geom_line()

ph_glorys_events
```

```{r}
th_glorys_clm <- th_glorys |>
  mhw_climatology() |>
  detect_events()

th_glorys_events <- th_glorys_clm |>
  calculate_events(funcs = event_funcs)

th_glorys_clm |>
  pivot_longer(
    c(temp, seas, thresh),
    names_to = "temp_type",
    values_to = "temp"
  ) |>
  ggplot(aes(x = t, y = temp, col = temp_type)) +
  geom_line()

th_glorys_events
```

## Comparing MHW event at GLORYS (2020-11-12)

```{r}
tibble(
  "data" = c("Observations", "GLORYS12v1"),
  "events" = c(nrow(ph_obs_daily_events),
               nrow(ph_glorys_events)),
  "mean_duration" = c(mean(ph_obs_daily_events$duration),
                      mean(ph_glorys_events$duration)),
  "mean_intensity" = c(mean(ph_obs_daily_events$intensity_mean),
                       mean(ph_glorys_events$intensity_mean)),
  "mean_cumulative_intensity" = c(mean(ph_obs_daily_events$intensity_cumulative),
                                  mean(ph_glorys_events$intensity_cumulative))
)

compare_event(
  clms = list(ph_obs_daily_clm, ph_glorys_clm),
  nums = c(3, 8),
  labels = c("Observations", "GLORYS12v1"),
  start_date = ymd("2020-10-23"),
  end_date = ymd("2020-12-02")
) +
  scale_x_datetime(date_labels = "%Y-%m") +
  theme(
    legend.position = "right",
    legend.background = element_rect(color = "white"),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(lwd = 1.2))) +
  labs(
    title = "",
    x = "Time"
  )
```

## Comparing at Taylor's Head (2020-12-06)

```{r}
tibble(
  "data" = c("Observations", "GLORYS12v1"),
  "events" = c(nrow(th_obs_daily_events),
               nrow(th_glorys_events)),
  "mean_duration" = c(mean(th_obs_daily_events$duration),
                      mean(th_glorys_events$duration)),
  "mean_intensity" = c(mean(th_obs_daily_events$intensity_mean),
                       mean(th_glorys_events$intensity_mean)),
  "mean_cumulative_intensity" = c(mean(th_obs_daily_events$intensity_cumulative),
                                  mean(th_glorys_events$intensity_cumulative))
)

compare_event(
  clms = list(th_obs_daily_clm, th_glorys_clm),
  nums = c(4, 9),
  labels = c("Observations", "GLORYS12v1"),
  start_date = ymd("2020-11-16"),
  end_date = ymd("2020-12-26")
) +
  scale_x_datetime(date_labels = "%Y-%m") +
  theme(
    legend.position = "right",
    legend.background = element_rect(color = "white"),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(lwd = 1.2))) +
  labs(
    title = "",
    x = "Time"
  )
```

## Trying to fit EVT models to daily data

```{r}
runtime <- 1

ph_glorys_evt_clm <- ph_glorys |>
  periodic_climatology2(
    filter_func = smooth.spline,
    filter_params = list("nknots" = 11),
    extract = "y",
    hourly = FALSE,
    padding = 30,
    thresh_quantile = 0.9
  )

thresholds <- seq(0, 3, length.out = 30)

for (i in 1:length(thresholds)) {
  
  ph_glorys_evt_clm |>
    mutate("thresh" = seas + thresholds[i]) |>
    detect_events(min_duration = 1, max_gap = runtime) |>
    calculate_events(funcs = event_funcs) |>
    nrow() |> print()
}
```

```{r}
runtime <- 1

th_glorys_evt_clm <- th_glorys |>
  periodic_climatology2(
    filter_func = smooth.spline,
    filter_params = list("nknots" = 9),
    extract = "y",
    hourly = FALSE,
    padding = 30,
    thresh_quantile = 0.9
  )

thresholds <- seq(0, 3, length.out = 30)

for (i in 1:length(thresholds)) {
  
  th_glorys_evt_clm |>
    mutate("thresh" = seas + thresholds[i]) |>
    detect_events(min_duration = 1, max_gap = runtime) |>
    calculate_events(funcs = event_funcs) |>
    nrow() |> print()
}
```

## Fitting EV models to GLORYS data over all years

```{r}
runtime <- 3

ph_glorys_full_clm <- daily_data$glorys |>
  filter(site == "Port L'Hebert") |>
  filter(year(t) < 2020) |>
  periodic_climatology2(
    filter_func = smooth.spline,
    filter_params = list("nknots" = 11),
    extract = "y",
    hourly = FALSE,
    padding = 30,
    thresh_quantile = 0.9
  )

ph_glorys_full_clm |>
  ggplot(aes(x = ind, y = temp)) +
  geom_line() +
  geom_line(data = filter(ph_glorys_full_clm, year(t) == 2000), aes(y = seas), col = "blue")

ph_glorys_full_anomalies <- na.omit(ph_glorys_full_clm$temp - ph_glorys_full_clm$seas)
fit_thresholds(ph_glorys_full_anomalies, range = c(1.2, 2.5), 
               nout = 30, alpha = 0.05, init = c(1,0)) +
  labs(
    title = "Threshold Stability Plot of Port L'Hebert GLORYS12v1"
  ) +
  theme_bw()

th_glorys_full_clm <- daily_data$glorys |>
  filter(site == "Taylor's Head") |>
  filter(year(t) < 2020) |>
  periodic_climatology(
    filter_func = smooth.spline,
    filter_params = list("nknots" = 11),
    extract = "y",
    hourly = FALSE,
    thresh_quantile = 0.9
  )

th_glorys_full_anomalies <- na.omit(th_glorys_full_clm$temp - th_glorys_full_clm$seas)
fit_thresholds(th_glorys_full_anomalies, range = c(1.2, 2.2), 
               nout = 30, alpha = 0.05, init = c(1,1)) +
  labs(
    title = "Threshold Stability Plot of Taylor's Head GLORYS12v1"
  ) +
  theme_bw()


# choose u = 1.6
# choose u = 1.7
```

```{r}
ph_glorys_full_clm <- ph_glorys_full_clm |>
  mutate("thresh" = seas + 1.4) |>
  detect_events(min_duration = 1, max_gap = runtime)
ph_glorys_full_events <- calculate_events(ph_glorys_full_clm, 
                                          funcs = event_funcs, 
                                          target_col = "exceedance")

th_glorys_full_clm <- th_glorys_full_clm |>
  mutate("thresh" = seas + 1.7) |>
  detect_events(min_duration = 1, max_gap = runtime)
th_glorys_full_events <- calculate_events(th_glorys_full_clm, 
                                          funcs = event_funcs, 
                                          target_col = "exceedance")

ph_glorys_full_exceedances <- ph_glorys_full_events$exceedance_max
ph_glorys_full_mle <- gp_mle(ph_glorys_full_exceedances, init = c(0,0))

th_glorys_full_exceedances <- th_glorys_full_events$exceedance_max
th_glorys_full_mle <- gp_mle(th_glorys_full_exceedances, init = c(0,0))

plot_gp_hist(ph_glorys_full_exceedances, ph_glorys_full_mle, pal[3], 0.7) +
  labs(
    x = "Exceedance [\u00B0C]",
    title = ""
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

plot_gp_hist(th_glorys_full_exceedances, th_glorys_full_mle, pal[3], 0.7) +
  labs(
    x = "Exceedance [\u00B0C]",
    title = ""
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )
```



## Testing Seasonal Cycle Definition

```{r}
original <- hourly |>
  filter(site == "Port L'Hebert") |>
  filter(type == "OBS") |>
  select(-type) |>
  periodic_climatology(
    filter_func = smooth.spline,
    filter_params = list(nknots = 9),
    extract = "y",
    hourly = TRUE,
    thresh_quantile = 0.9
  ) |>
  mutate("type" = "original")

new <- hourly |>
  filter(site == "Port L'Hebert") |>
  filter(type == "OBS") |>
  select(-type) |>
  periodic_climatology2(
    filter_func = smooth.spline,
    filter_params = list(nknots = 11),
    extract = "y",
    hourly = TRUE,
    padding = 30*24,
    thresh_quantile = 0.9
  ) |>
  mutate("type" = "new")

bind_rows(original, new) |>
  ggplot(aes(x = t, y = temp)) +
  geom_line(data = original) +
  geom_line(aes(y = seas, col = type))

bind_rows(original, new) |>
  filter(year(t) == 2019) |>
  ggplot(aes(x = t, y = temp)) +
  geom_line(data = filter(original, year(t) == 2019)) +
  geom_line(aes(y = seas, col = type))
```


