---
title: "Frequency Validation 2"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Code setup

Loading libraries and functions, reading and preprocessing data.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(heatwaveR)
library(ncdf4)
library(R.matlab)
#library(patchwork)
#library(logKDE)
```

```{r}
read_data = FALSE
process = "sst"
sites <- c("Port l'Hebert", "Taylor's Head Shallow")
site_labels <- c("Port L'Hebert", "Taylor's Head")
pal <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442")
source("setup.R")
```

# Reading CMIP6

```{r}
ph_cmip6 <- readCMIP6("../data/Port_LHebert/cmip6_2000_to_2024.mat")
th_cmip6 <- readCMIP6("../data/TaylorHead/cmip6_2000_to_2024.mat")

ph_cmip6 <- ph_cmip6 |>
  mutate("site" = "Port L'Hebert")
th_cmip6 <- th_cmip6 |>
  mutate("site" = "Taylor's Head")
cmip6 <- bind_rows(ph_cmip6, th_cmip6) |>
  mutate("type" = "CMIP6") |>
  select("t" = time, site, type, temp)
```

# Processing Data

Filtering between May 28th, 2018 to March 20th, 2020.

```{r}
hourly <- hourly_dataframe |>
  filter(between(time, dates$freq[1], dates$freq[2])) |>
  select(-GLORYS) |>
  rename("t" = time) |>
  pivot_longer(
    c(OBS, FVCOM, FVCOM2),
    names_to = "type",
    values_to = "temp"
  ) |>
  remove_missing_times()

daily <- daily_dataframe |>
  filter(between(time, dates$freq[1], dates$freq[2])) |>
  rename("t" = time) |>
  pivot_longer(
    c(OBS, FVCOM, FVCOM2, GLORYS),
    names_to = "type",
    values_to = "temp"
  ) |>
  remove_missing_times()

monthly <- daily |>
  bind_rows(cmip6) |>
  filter(between(t, dates$freq[1], dates$freq[2])) |>
  mutate(
    "year" = year(t),
    "month" = month(t)
  ) |>
  group_by(site, type, year, month) |>
  summarize(
    "temp" = mean(temp),
    .groups = "drop"
  ) |>
  mutate(
    "t" = make_date(year, month),
  ) |>
  select(t, site, type, temp)
```

## Decomposing Temeprature Time Series

```{r}
hourly_filtered <- hourly |>
  group_by(site, type) |>
  mutate(
    "seas" = smooth.spline(temp, nknots = 16)$y,
    "anomaly" = temp - seas,
    "met" = freq_filter(anomaly, 1, 1/48, type = "lowpass"),
    "tidal" = anomaly - met
  ) |>
  ungroup()

daily_filtered <- daily |>
  group_by(site, type) |>
  mutate(
    "seas" = smooth.spline(temp, nknots = 16)$y,
    "anomaly" = temp - seas,
    "met" = freq_filter(anomaly, 1, 1/2, type = "lowpass"),
    "tidal" = anomaly - met
  ) |>
  ungroup()
```

# Plotting Auto-correlation Functions

```{r}
hourly_filtered |>
  group_by(site, type) |>
  reframe(
    "lag" = acf(anomaly, lag.max = 500, type = "correlation", plot = FALSE)$lag,
    "acf" = acf(anomaly, lag.max = 500, type = "correlation", plot = FALSE)$acf
  ) |>
  mutate(type = factor(type, levels = c("OBS", "FVCOM", "FVCOM2"))) |>
  ggplot(aes(x = lag, y = acf)) +
  geom_hline(aes(yintercept = 0)) +
  geom_segment(aes(xend = lag, yend = 0)) +
  facet_grid(type ~ site, 
             labeller = as_labeller(c("FVCOM" = "FVCOMv1",
                                      "FVCOM2" = "FVCOMv2",
                                      "OBS" = "Observations",
                                      "Port L'Hebert" = "Port L'Hebert",
                                      "Taylor's Head" = "Taylor's Head"))) +
  theme_bw() +
  labs(
    title = "Auto-correlation Functions of the Anomalies",
    x = "Lag",
    y = "Auto-correlation"
  )
```

# Plotting Cross-correlation Functions

```{r}
hourly_filtered |>
  select(t, site, type, anomaly) |>
  pivot_wider(
    names_from = type,
    values_from = anomaly
  ) |>
  pivot_longer(
    c(FVCOM, FVCOM2),
    names_to = "type",
    values_to = "anomaly"
  ) |>
  group_by(site, type) |>
  reframe(
    "lag" = ccf(OBS, anomaly, lag.max = 250, type = "correlation", plot = FALSE)$lag[,1,1],
    "ccf" = ccf(OBS, anomaly, lag.max = 250, type = "correlation", plot = FALSE)$acf[,1,1]
  ) |>
  mutate(type = factor(type, levels = c("FVCOM", "FVCOM2"))) |>
  ggplot(aes(x = lag, y = ccf)) +
  geom_hline(aes(yintercept = 0)) +
  geom_segment(aes(xend = lag, yend = 0)) +
  facet_grid(type ~ site, 
             labeller = as_labeller(c("FVCOM" = "FVCOMv1",
                                      "FVCOM2" = "FVCOMv2",
                                      "Port L'Hebert" = "Port L'Hebert",
                                      "Taylor's Head" = "Taylor's Head"))) +
  theme_bw() +
  labs(
    title = "Cross-correlation of the Anomalies with Respect to the Observations",
    x = "Lag",
    y = "Cross-correlation"
  )
```

# Plotting Power Spectrums

```{r}
hourly_spectrum <- hourly_filtered |>
  group_by(site, type) |>
  reframe(
    "freq" = spec.pgram(anomaly, spans = 21, plot = FALSE)$freq,
    "power" = spec.pgram(anomaly, spans = 21, plot = FALSE)$spec
  )

glorys_spectrum <- daily_filtered |>
  filter(type == "GLORYS") |>
  group_by(site) |>
  reframe(
    "freq" = spec.pgram(anomaly, spans = 21, plot = FALSE)$freq,
    "power" = spec.pgram(anomaly, spans = 21, plot = FALSE)$spec
  ) |>
  mutate("freq" = freq / 24) |>
  mutate("type" = "GLORYS")

bind_rows(hourly_spectrum, glorys_spectrum) |>
  filter(freq < 0.1) |>
  mutate(type = factor(type, levels = c("OBS", "FVCOM", "FVCOM2", "GLORYS"))) |>
  ggplot(aes(x = freq, y = log(power))) +
  geom_line() +
  geom_vline(aes(xintercept = 1 / 25.82), lty = 2, lwd = 0.7, col = pal[3]) +
  geom_vline(aes(xintercept = 1 / 24), lty = 2, lwd = 0.7, col = pal[1]) +
  geom_vline(aes(xintercept = 1 / 12.42), lty = 2, lwd = 0.7, col = pal[4]) +
  geom_vline(aes(xintercept = 1 / 12), lty = 2, lwd = 0.7, col = pal[2]) +
  facet_grid(type ~ site, 
             labeller = as_labeller(c("FVCOM" = "FVCOMv1",
                                      "FVCOM2" = "FVCOMv2",
                                      "OBS" = "Observations",
                                      "GLORYS" = "GLORYS12v1",
                                      "Port L'Hebert" = "Port L'Hebert",
                                      "Taylor's Head" = "Taylor's Head"))) +
  scale_x_continuous(
    breaks = c(0, 0.025, 0.050, 0.075, 0.1),
    labels = c("0", "0.025", "0.05", "0.075", "0.1")
  ) +
  labs(
    title = "",
    x = "Frequency (1/hour)",
    y = "Log power"
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12),
  )
```

## Estimating Slope of the High Frequencies

```{r}
hourly_filtered |>
  group_by(site, type) |>
  reframe(
    "freq" = spec.pgram(anomaly, spans = 21, plot = FALSE)$freq,
    "power" = spec.pgram(anomaly, spans = 21, plot = FALSE)$spec
  ) |>
  filter(freq > 1/6) |>
  mutate(type = factor(type, levels = c("OBS", "FVCOM", "FVCOM2"))) |>
  ggplot(aes(x = log(freq), y = log(power))) +
  geom_line() +
  geom_smooth(method = "lm", formula = y ~ x, col = "red") +
  facet_grid(type ~ site, 
             labeller = as_labeller(c("FVCOM" = "FVCOMv1",
                                      "FVCOM2" = "FVCOMv2",
                                      "OBS" = "Observations",
                                      "Port L'Hebert" = "Port L'Hebert",
                                      "Taylor's Head" = "Taylor's Head"))) +
  labs(
    title = "",
    x = "Log frequency",
    y = "Log power"
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12),
  )
```

```{r}
hourly_filtered |>
  group_by(site, type) |>
  reframe(
    "freq" = spec.pgram(anomaly, spans = 21, plot = FALSE)$freq,
    "power" = spec.pgram(anomaly, spans = 21, plot = FALSE)$spec
  ) |>
  filter(freq > 1/6) |>
  mutate(type = factor(type, levels = c("OBS", "FVCOM", "FVCOM2"))) |>
  group_by(site, type) |>
  summarize(
    "slope" = get_slope(log(freq), log(power)),
    .groups = "drop"
  )
```

# Time Series Decomposition

## Port L'Hebert FVCOMv1

```{r}
hourly_filtered |>
  filter(site == "Port L'Hebert") |>
  filter(type %in% c("OBS", "FVCOM")) |>
  pivot_longer(
    c(seas, met, tidal),
    names_to = "temp_type",
    values_to = "temp_value"
  ) |>
  mutate(
    type = factor(type, levels = c("OBS", "FVCOM")),
    temp_type = factor(temp_type, levels = c("seas", "met", "tidal"))
  ) |>
  ggplot(aes(x = t, y = temp_value, col = type)) +
  geom_line(lwd = 0.7) +
  facet_wrap(~temp_type, ncol = 1, scales = "free",
             labeller = as_labeller(c("seas" = "Seasonal",
                                      "met" = "Meteorological",
                                      "tidal" = "Tidal"))) +
  scale_color_manual(
    name = "",
    values = c("black", "blue"),
    breaks = c("OBS", "FVCOM"),
    labels = c("Observations", "FVCOMv1")
  ) +
  labs(
    title = "",
    x = "Time",
    y = "Temperature [\u00B0C]"
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    legend.key.width = unit(2, "cm"),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(lwd = 1.2)))
```

## Port L'Hebert FVCOMv2

```{r}
hourly_filtered |>
  filter(site == "Port L'Hebert") |>
  filter(type %in% c("OBS", "FVCOM2")) |>
  pivot_longer(
    c(seas, met, tidal),
    names_to = "temp_type",
    values_to = "temp_value"
  ) |>
  mutate(
    type = factor(type, levels = c("OBS", "FVCOM2")),
    temp_type = factor(temp_type, levels = c("seas", "met", "tidal"))
  ) |>
  ggplot(aes(x = t, y = temp_value, col = type)) +
  geom_line(lwd = 0.7) +
  facet_wrap(~temp_type, ncol = 1, scales = "free",
             labeller = as_labeller(c("seas" = "Seasonal",
                                      "met" = "Meteorological",
                                      "tidal" = "Tidal"))) +
  scale_color_manual(
    name = "",
    values = c("black", "blue"),
    breaks = c("OBS", "FVCOM2"),
    labels = c("Observations", "FVCOMv2")
  ) +
  labs(
    title = "",
    x = "Time",
    y = "Temperature [\u00B0C]"
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    legend.key.width = unit(2, "cm"),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(lwd = 1.2)))
```

## Taylor's Head FVCOMv1

```{r}
hourly_filtered |>
  filter(site == "Taylor's Head") |>
  filter(type %in% c("OBS", "FVCOM")) |>
  pivot_longer(
    c(seas, met, tidal),
    names_to = "temp_type",
    values_to = "temp_value"
  ) |>
  mutate(
    type = factor(type, levels = c("OBS", "FVCOM")),
    temp_type = factor(temp_type, levels = c("seas", "met", "tidal"))
  ) |>
  ggplot(aes(x = t, y = temp_value, col = type)) +
  geom_line(lwd = 0.7) +
  facet_wrap(~temp_type, ncol = 1, scales = "free",
             labeller = as_labeller(c("seas" = "Seasonal",
                                      "met" = "Meteorological",
                                      "tidal" = "Tidal"))) +
  scale_color_manual(
    name = "",
    values = c("black", "blue"),
    breaks = c("OBS", "FVCOM"),
    labels = c("Observations", "FVCOMv1")
  ) +
  labs(
    title = "",
    x = "Time",
    y = "Temperature [\u00B0C]"
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    legend.key.width = unit(2, "cm"),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(lwd = 1.2)))
```

## Taylor's Head FVCOMv2

```{r}
hourly_filtered |>
  filter(site == "Taylor's Head") |>
  filter(type %in% c("OBS", "FVCOM2")) |>
  pivot_longer(
    c(seas, met, tidal),
    names_to = "temp_type",
    values_to = "temp_value"
  ) |>
  mutate(
    type = factor(type, levels = c("OBS", "FVCOM2")),
    temp_type = factor(temp_type, levels = c("seas", "met", "tidal"))
  ) |>
  ggplot(aes(x = t, y = temp_value, col = type)) +
  geom_line(lwd = 0.7) +
  facet_wrap(~temp_type, ncol = 1, scales = "free",
             labeller = as_labeller(c("seas" = "Seasonal",
                                      "met" = "Meteorological",
                                      "tidal" = "Tidal"))) +
  scale_color_manual(
    name = "",
    values = c("black", "blue"),
    breaks = c("OBS", "FVCOM2"),
    labels = c("Observations", "FVCOMv2")
  ) +
  labs(
    title = "",
    x = "Time",
    y = "Temperature [\u00B0C]"
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    legend.key.width = unit(2, "cm"),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(lwd = 1.2)))
```

# Calculating Metrics for Specific Bands

## Port L'Hebert

```{r}
ph_hourly_filtered_wide  <- hourly_filtered |>
  filter(site == "Port L'Hebert") |>
  pivot_longer(
    c(temp, seas, met, tidal),
    names_to = "temp_type",
    values_to = "temp"
  ) |>
  unite("type", type, temp_type, sep = "_") |>
  select(t, type, temp) |>
  pivot_wider(
    names_from = type,
    values_from = temp
  )

bind_cols(
  apply_metrics(
    ph_hourly_filtered_wide$OBS_temp,
    ph_hourly_filtered_wide$FVCOM_temp,
    pred_label = "FVCOMv1 Temp",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ),
  apply_metrics(
    ph_hourly_filtered_wide$OBS_seas,
    ph_hourly_filtered_wide$FVCOM_seas,
    pred_label = "FVCOMv1 Seas",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric),
  apply_metrics(
    ph_hourly_filtered_wide$OBS_met,
    ph_hourly_filtered_wide$FVCOM_met,
    pred_label = "FVCOMv1 Met",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric),
  apply_metrics(
    ph_hourly_filtered_wide$OBS_tidal,
    ph_hourly_filtered_wide$FVCOM_tidal,
    pred_label = "FVCOMv1 Tidal",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric)
) |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

bind_cols(
  apply_metrics(
    ph_hourly_filtered_wide$OBS_temp,
    ph_hourly_filtered_wide$FVCOM2_temp,
    pred_label = "FVCOMv2 Temp",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ),
  apply_metrics(
    ph_hourly_filtered_wide$OBS_seas,
    ph_hourly_filtered_wide$FVCOM2_seas,
    pred_label = "FVCOMv2 Seas",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric),
  apply_metrics(
    ph_hourly_filtered_wide$OBS_met,
    ph_hourly_filtered_wide$FVCOM2_met,
    pred_label = "FVCOMv2 Met",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric),
  apply_metrics(
    ph_hourly_filtered_wide$OBS_tidal,
    ph_hourly_filtered_wide$FVCOM2_tidal,
    pred_label = "FVCOMv2 Tidal",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric)
) |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))
```

## Taylor's Head

```{r}
th_hourly_filtered_wide  <- hourly_filtered |>
  filter(site == "Taylor's Head") |>
  pivot_longer(
    c(temp, seas, met, tidal),
    names_to = "temp_type",
    values_to = "temp"
  ) |>
  unite("type", type, temp_type, sep = "_") |>
  select(t, type, temp) |>
  pivot_wider(
    names_from = type,
    values_from = temp
  )

bind_cols(
  apply_metrics(
    th_hourly_filtered_wide$OBS_temp,
    th_hourly_filtered_wide$FVCOM_temp,
    pred_label = "FVCOMv1 Temp",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ),
  apply_metrics(
    th_hourly_filtered_wide$OBS_seas,
    th_hourly_filtered_wide$FVCOM_seas,
    pred_label = "FVCOMv1 Seas",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric),
  apply_metrics(
    th_hourly_filtered_wide$OBS_met,
    th_hourly_filtered_wide$FVCOM_met,
    pred_label = "FVCOMv1 Met",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric),
  apply_metrics(
    th_hourly_filtered_wide$OBS_tidal,
    th_hourly_filtered_wide$FVCOM_tidal,
    pred_label = "FVCOMv1 Tidal",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric)
) |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

bind_cols(
  apply_metrics(
    th_hourly_filtered_wide$OBS_temp,
    th_hourly_filtered_wide$FVCOM2_temp,
    pred_label = "FVCOMv2 Temp",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ),
  apply_metrics(
    th_hourly_filtered_wide$OBS_seas,
    th_hourly_filtered_wide$FVCOM2_seas,
    pred_label = "FVCOMv2 Seas",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric),
  apply_metrics(
    th_hourly_filtered_wide$OBS_met,
    th_hourly_filtered_wide$FVCOM2_met,
    pred_label = "FVCOMv2 Met",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric),
  apply_metrics(
    th_hourly_filtered_wide$OBS_tidal,
    th_hourly_filtered_wide$FVCOM2_tidal,
    pred_label = "FVCOMv2 Tidal",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric)
) |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))
```

# Daily GLORYS12v1 decomposition

## Port L'Hebert

```{r}
daily_filtered |>
  filter(site == "Port L'Hebert") |>
  filter(type %in% c("OBS", "GLORYS")) |>
  pivot_longer(
    c(seas, met),
    names_to = "temp_type",
    values_to = "temp_value"
  ) |>
  mutate(
    type = factor(type, levels = c("OBS", "GLORYS")),
    temp_type = factor(temp_type, levels = c("seas", "met"))
  ) |>
  ggplot(aes(x = t, y = temp_value, col = type)) +
  geom_line(lwd = 0.7) +
  facet_wrap(~temp_type, ncol = 1, scales = "free",
             labeller = as_labeller(c("seas" = "Seasonal",
                                      "met" = "Meteorological"))) +
  scale_color_manual(
    name = "",
    values = c("black", "blue"),
    breaks = c("OBS", "GLORYS"),
    labels = c("Observations", "GLORYS12v1")
  ) +
  labs(
    title = "",
    x = "Time",
    y = "Temperature [\u00B0C]"
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    legend.key.width = unit(2, "cm"),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(lwd = 1.2)))
```

## Taylor's Head

```{r}
daily_filtered |>
  filter(site == "Taylor's Head") |>
  filter(type %in% c("OBS", "GLORYS")) |>
  pivot_longer(
    c(seas, met),
    names_to = "temp_type",
    values_to = "temp_value"
  ) |>
  mutate(
    type = factor(type, levels = c("OBS", "GLORYS")),
    temp_type = factor(temp_type, levels = c("seas", "met"))
  ) |>
  ggplot(aes(x = t, y = temp_value, col = type)) +
  geom_line(lwd = 0.7) +
  facet_wrap(~temp_type, ncol = 1, scales = "free",
             labeller = as_labeller(c("seas" = "Seasonal",
                                      "met" = "Meteorological"))) +
  scale_color_manual(
    name = "",
    values = c("black", "blue"),
    breaks = c("OBS", "GLORYS"),
    labels = c("Observations", "GLORYS12v1")
  ) +
  labs(
    title = "",
    x = "Time",
    y = "Temperature [\u00B0C]"
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    legend.key.width = unit(2, "cm"),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(lwd = 1.2)))
```

## Calculating Daily Metrics

```{r}
ph_daily_filtered_wide  <- daily_filtered |>
  filter(site == "Port L'Hebert") |>
  pivot_longer(
    c(temp, seas, met),
    names_to = "temp_type",
    values_to = "temp"
  ) |>
  unite("type", type, temp_type, sep = "_") |>
  select(t, type, temp) |>
  pivot_wider(
    names_from = type,
    values_from = temp
  )

th_daily_filtered_wide  <- daily_filtered |>
  filter(site == "Taylor's Head") |>
  pivot_longer(
    c(temp, seas, met),
    names_to = "temp_type",
    values_to = "temp"
  ) |>
  unite("type", type, temp_type, sep = "_") |>
  select(t, type, temp) |>
  pivot_wider(
    names_from = type,
    values_from = temp
  )

bind_cols(
  apply_metrics(
    ph_daily_filtered_wide$OBS_temp,
    ph_daily_filtered_wide$GLORYS_temp,
    pred_label = "GLORYS Temp",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ),
  apply_metrics(
    ph_daily_filtered_wide$OBS_seas,
    ph_daily_filtered_wide$GLORYS_seas,
    pred_label = "GLORYS Seas",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric),
  apply_metrics(
    ph_daily_filtered_wide$OBS_met,
    ph_daily_filtered_wide$GLORYS_met,
    pred_label = "GLORYS Met",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric)
) |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

bind_cols(
  apply_metrics(
    th_daily_filtered_wide$OBS_temp,
    th_daily_filtered_wide$GLORYS_temp,
    pred_label = "GLORYS Temp",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ),
  apply_metrics(
    th_daily_filtered_wide$OBS_seas,
    th_daily_filtered_wide$GLORYS_seas,
    pred_label = "GLORYS Seas",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric),
  apply_metrics(
    th_daily_filtered_wide$OBS_met,
    th_daily_filtered_wide$GLORYS_met,
    pred_label = "GLORYS Met",
    metric_funcs = list(bias, rmse, sig_star, cor, willmott, nse),
    metric_labels = c("B", "RMSE", "sig_star", "R", "WS", "NSE")
  ) |> select(-metric)
) |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))
```


