---
title: "Time-Dependent Validation"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Code setup

Loading libraries and functions, reading and preprocessing data.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(heatwaveR)
library(ncdf4)
library(R.matlab)
#library(patchwork)
#library(logKDE)
```

```{r}
read_data = FALSE
process = "sst"
sites <- c("Port l'Hebert", "Taylor's Head Shallow")
site_labels <- c("Port L'Hebert", "Taylor's Head")
pal <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442")
source("setup.R")
```

# Reading CMIP6

```{r}
ph_cmip6 <- readCMIP6("../data/Port_LHebert/cmip6_2000_to_2024.mat")
th_cmip6 <- readCMIP6("../data/TaylorHead/cmip6_2000_to_2024.mat")

ph_cmip6 <- ph_cmip6 |>
  mutate("site" = "Port L'Hebert")
th_cmip6 <- th_cmip6 |>
  mutate("site" = "Taylor's Head")
cmip6 <- bind_rows(ph_cmip6, th_cmip6) |>
  mutate("type" = "CMIP6") |>
  select("t" = time, site, type, temp)
```

# Processing Data

Filtering between May 28th, 2018 to March 20th, 2020.

```{r}
hourly <- hourly_dataframe |>
  filter(between(time, dates$freq[1], dates$freq[2])) |>
  select(-GLORYS) |>
  rename("t" = time) |>
  pivot_longer(
    c(OBS, FVCOM, FVCOM2),
    names_to = "type",
    values_to = "temp"
  ) |>
  remove_missing_times()

daily <- daily_dataframe |>
  filter(between(time, dates$freq[1], dates$freq[2])) |>
  rename("t" = time) |>
  pivot_longer(
    c(OBS, FVCOM, FVCOM2, GLORYS),
    names_to = "type",
    values_to = "temp"
  ) |>
  remove_missing_times()

monthly <- daily |>
  bind_rows(cmip6) |>
  filter(between(t, dates$freq[1], dates$freq[2])) |>
  mutate(
    "year" = year(t),
    "month" = month(t)
  ) |>
  group_by(site, type, year, month) |>
  summarize(
    "temp" = mean(temp),
    .groups = "drop"
  ) |>
  mutate(
    "t" = make_date(year, month),
  ) |>
  select(t, site, type, temp)
```

## Decomposing Temeprature Time Series

```{r}
hourly_filtered <- hourly |>
  group_by(site, type) |>
  mutate(
    "seas" = smooth.spline(temp, nknots = 16)$y,
    "anomaly" = temp - seas,
    "met" = freq_filter(anomaly, 1, 1/48, type = "lowpass"),
    "tidal" = anomaly - met
  ) |>
  ungroup()

daily_filtered <- daily |>
  group_by(site, type) |>
  mutate(
    "seas" = smooth.spline(temp, nknots = 16)$y,
    "anomaly" = temp - seas,
    "met" = freq_filter(anomaly, 1, 1/2, type = "lowpass"),
    "tidal" = anomaly - met
  ) |>
  ungroup()
```

# Calculating Evolutionary Metrics

## Constructing Dataframe

```{r}
evolutionary_filtered <- bind_rows(
  hourly_filtered |>
    select(t, site, type, temp) |>
    pivot_wider(
      names_from = type,
      values_from = temp
    ) |>
    pivot_longer(
      c(FVCOM, FVCOM2),
      names_to = "type",
      values_to = "temp"
    ),
  daily_filtered |>
    filter(type %in% c("OBS", "GLORYS")) |>
    select(t, site, type, temp) |>
    mutate(t = as_datetime(t)) |>
    pivot_wider(
      names_from = type,
      values_from = temp
    ) |>
    pivot_longer(
      c(GLORYS),
      names_to = "type",
      values_to = "temp"
    )
) |>
  mutate(type = factor(type, levels = c("FVCOM", "FVCOM2", "GLORYS")))
```

# Overall Bias (FVCOM models)

```{r}
evolutionary_filtered |>
  filter(type != "GLORYS") |>
  group_by(site, type) |>
  reframe(
    "time" = evolutionary_metrics(t, OBS, temp, bias)$t,
    "metric" = evolutionary_metrics(t, OBS, temp, bias)$metric
  ) |>
  mutate("month" = factor(month(time) %in% c(5, 6, 7, 8, 9, 10))) |>
  ggplot(aes(x = time, y = metric)) +
  geom_line() +
  geom_hline(aes(yintercept = 0), lty = 2) +
  geom_point(aes(col = month)) +
  scale_color_manual(
    name = "",
    values = c("blue", "red"),
    breaks = c("FALSE", "TRUE"),
    labels = c("Winter", "Summer")
  ) +
  facet_grid(type ~ site, 
             labeller = as_labeller(c("FVCOM" = "FVCOMv1",
                                      "FVCOM2" = "FVCOMv2",
                                      "Port L'Hebert" = "Port L'Hebert",
                                      "Taylor's Head" = "Taylor's Head"))) +
  theme_bw() +
  labs(
    title = "",
    x = "Time",
    y = "B"
  ) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )
```

## RMSE (FVCOM models)

```{r}
evolutionary_filtered |>
  filter(type != "GLORYS") |>
  group_by(site, type) |>
  reframe(
    "time" = evolutionary_metrics(t, OBS, temp, rmse)$t,
    "metric" = evolutionary_metrics(t, OBS, temp, rmse)$metric
  ) |>
  mutate("month" = factor(month(time) %in% c(5, 6, 7, 8, 9, 10))) |>
  ggplot(aes(x = time, y = metric)) +
  geom_line() +
  geom_point(aes(col = month)) +
  scale_color_manual(
    name = "",
    values = c("blue", "red"),
    breaks = c("FALSE", "TRUE"),
    labels = c("Winter", "Summer")
  ) +
  facet_grid(type ~ site, 
             labeller = as_labeller(c("FVCOM" = "FVCOMv1",
                                      "FVCOM2" = "FVCOMv2",
                                      "Port L'Hebert" = "Port L'Hebert",
                                      "Taylor's Head" = "Taylor's Head"))) +
  theme_bw() +
  labs(
    title = "",
    x = "Time",
    y = "RMSE"
  ) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )
```

## Correlation Coefficient (FVCOM models)

```{r}
evolutionary_filtered |>
  filter(type != "GLORYS") |>
  group_by(site, type) |>
  reframe(
    "time" = evolutionary_metrics(t, OBS, temp, r)$t,
    "metric" = evolutionary_metrics(t, OBS, temp, r)$metric
  ) |>
  mutate("month" = factor(month(time) %in% c(5, 6, 7, 8, 9, 10))) |>
  ggplot(aes(x = time, y = metric)) +
  geom_line() +
  geom_point(aes(col = month)) +
  scale_color_manual(
    name = "",
    values = c("blue", "red"),
    breaks = c("FALSE", "TRUE"),
    labels = c("Winter", "Summer")
  ) +
  facet_grid(type ~ site, 
             labeller = as_labeller(c("FVCOM" = "FVCOMv1",
                                      "FVCOM2" = "FVCOMv2",
                                      "Port L'Hebert" = "Port L'Hebert",
                                      "Taylor's Head" = "Taylor's Head"))) +
  theme_bw() +
  labs(
    title = "",
    x = "Time",
    y = "R"
  ) +
  theme(
    legend.position = "none"
  ) +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.9, 0.15),
    legend.background = element_blank(),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(size = 3)))
```

## Willmott Skill (FVCOM models)

```{r}
evolutionary_filtered |>
  filter(type != "GLORYS") |>
  group_by(site, type) |>
  reframe(
    "time" = evolutionary_metrics(t, OBS, temp, willmott)$t,
    "metric" = evolutionary_metrics(t, OBS, temp, willmott)$metric
  ) |>
  mutate("month" = factor(month(time) %in% c(5, 6, 7, 8, 9, 10))) |>
  ggplot(aes(x = time, y = metric)) +
  geom_line() +
  geom_point(aes(col = month)) +
  scale_color_manual(
    name = "",
    values = c("blue", "red"),
    breaks = c("FALSE", "TRUE"),
    labels = c("Winter", "Summer")
  ) +
  facet_grid(type ~ site, 
             labeller = as_labeller(c("FVCOM" = "FVCOMv1",
                                      "FVCOM2" = "FVCOMv2",
                                      "Port L'Hebert" = "Port L'Hebert",
                                      "Taylor's Head" = "Taylor's Head"))) +
  theme_bw() +
  labs(
    title = "",
    x = "Time",
    y = "WS"
  ) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )
```

## Evolutionary RMSE for Specific Bands (Port L'Hebert) 

```{r}
hourly_filtered |>
  filter(site == "Port L'Hebert") |>
  select(t, type, seas, met, tidal) |>
  pivot_longer(
    c(seas, met, tidal),
    names_to = "band",
    values_to = "temp"
  ) |>
  pivot_wider(
    names_from = type,
    values_from = temp
  ) |>
  pivot_longer(
    c(FVCOM, FVCOM2),
    names_to = "type",
    values_to = "temp"
  ) |>
  group_by(type, band) |>
  reframe(
    "time" = evolutionary_metrics(t, OBS, temp, rmse)$t,
    "rmse" = evolutionary_metrics(t, OBS, temp, rmse)$metric
  ) |>
  mutate(
    "month" = factor(month(time) %in% c(5, 6, 7, 8, 9, 10))
  ) |>
  mutate(type = factor(type, levels = c("FVCOM", "FVCOM2"))) |>
  mutate(band = factor(band, levels = c("seas", "met", "tidal"))) |>
  ggplot(aes(x = time, y = rmse)) +
  geom_line() +
  geom_point(aes(col = month)) +
  scale_color_manual(
    name = "",
    values = c("blue", "red"),
    breaks = c("FALSE", "TRUE"),
    labels = c("Winter", "Summer")
  ) +
  facet_grid(band ~ type, 
             labeller = as_labeller(c("FVCOM" = "FVCOMv1",
                                      "FVCOM2" = "FVCOMv2",
                                      "seas" = "Seasonal",
                                      "met" = "Meteorological",
                                      "tidal" = "Tidal"))) +
  labs(
    title = "",
    x = "Time",
    y = "RMSE"
  ) +
  theme_bw() +
  theme(
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(size = 2)))
```

## Evolutionary RMSE for Specific Bands (Taylor's Head)

```{r}
hourly_filtered |>
  filter(site == "Taylor's Head") |>
  select(t, type, seas, met, tidal) |>
  pivot_longer(
    c(seas, met, tidal),
    names_to = "band",
    values_to = "temp"
  ) |>
  pivot_wider(
    names_from = type,
    values_from = temp
  ) |>
  pivot_longer(
    c(FVCOM, FVCOM2),
    names_to = "type",
    values_to = "temp"
  ) |>
  group_by(type, band) |>
  reframe(
    "time" = evolutionary_metrics(t, OBS, temp, rmse)$t,
    "rmse" = evolutionary_metrics(t, OBS, temp, rmse)$metric
  ) |>
  mutate(
    "month" = factor(month(time) %in% c(5, 6, 7, 8, 9, 10))
  ) |>
  mutate(type = factor(type, levels = c("FVCOM", "FVCOM2"))) |>
  mutate(band = factor(band, levels = c("seas", "met", "tidal"))) |>
  ggplot(aes(x = time, y = rmse)) +
  geom_line() +
  geom_point(aes(col = month)) +
  scale_color_manual(
    name = "",
    values = c("blue", "red"),
    breaks = c("FALSE", "TRUE"),
    labels = c("Winter", "Summer")
  ) +
  facet_grid(band ~ type, 
             labeller = as_labeller(c("FVCOM" = "FVCOMv1",
                                      "FVCOM2" = "FVCOMv2",
                                      "seas" = "Seasonal",
                                      "met" = "Meteorological",
                                      "tidal" = "Tidal"))) +
  labs(
    title = "",
    x = "Time",
    y = "RMSE"
  ) +
  theme_bw() +
  theme(
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(size = 2)))
```

## Evolutionary Metrics for GLORYS12v1

```{r}
evolutionary_filtered |>
  filter(type == "GLORYS") |>
  group_by(site) |>
  reframe(
    "time" = evolutionary_metrics(t, OBS, temp, bias, width = 30, overlap = 15)$t,
    "bias" = evolutionary_metrics(t, OBS, temp, bias, width = 30, overlap = 15)$metric,
    "rmse" = evolutionary_metrics(t, OBS, temp, rmse, width = 30, overlap = 15)$metric,
    "willmott" = evolutionary_metrics(t, OBS, temp, willmott, width = 30, overlap = 15)$metric,
    "cor" = evolutionary_metrics(t, OBS, temp, r, width = 30, overlap = 15)$metric,
  ) |>
  pivot_longer(
    c(bias, rmse, willmott, cor),
    names_to = "metric",
    values_to = "value"
  ) |>
  mutate("month" = factor(month(time) %in% c(5, 6, 7, 8, 9, 10))) |>
  ggplot(aes(x = time, y = value)) +
  geom_line() +
  geom_point(aes(col = month)) +
  scale_color_manual(
    name = "",
    values = c("blue", "red"),
    breaks = c("FALSE", "TRUE"),
    labels = c("Winter", "Summer")
  ) +
  facet_grid(metric ~ site, 
             labeller = as_labeller(c("bias" = "B",
                                      "rmse" = "RMSE",
                                      "willmott" = "WS",
                                      "cor" = "R",
                                      "Port L'Hebert" = "Port L'Hebert",
                                      "Taylor's Head" = "Taylor's Head")),
             scales = "free") +
  theme_bw() +
  labs(
    title = "",
    x = "Time",
    y = "Metric value"
  ) +
  theme(
    legend.position = "right"
  ) +
  theme(
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  ) +
  guides(col = guide_legend(override.aes = list(size = 2)))
```

# Spectrograms

```{r}
hourly_filtered |>
  group_by(site, type) |>
  reframe(
    "time" = evolutionary_spectrum2(t, anomaly)$t,
    "freq" = evolutionary_spectrum2(t, anomaly)$freq,
    "power" = evolutionary_spectrum2(t, anomaly)$power
  ) |>
  filter(freq < 1/6) |>
  mutate(type = factor(type, levels = c("OBS", "FVCOM", "FVCOM2"))) |>
  ggplot(aes(x = time, y = freq, fill = log(power))) +
  geom_tile() +
  scale_fill_gradientn(colors = c("darkblue", "blue", "lightblue1",
                                    "green", "yellow", "red", "darkred"),
                       name = "Log power") +
  facet_grid(type ~ site, 
             labeller = as_labeller(c("FVCOM" = "FVCOMv1",
                                      "FVCOM2" = "FVCOMv2",
                                      "OBS" = "Observations",
                                      "Port L'Hebert" = "Port L'Hebert",
                                      "Taylor's Head" = "Taylor's Head"))) +
  labs(
    x = "Time",
    y = "Frequency (1/hour)",
    title = ""
  ) +
  theme_bw() +
  theme(
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )
```