---
title: "Traditional Validation"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Code setup

Loading libraries and functions, reading and preprocessing data.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(heatwaveR)
library(ncdf4)
library(R.matlab)
#library(patchwork)
#library(logKDE)
```

```{r}
read_data = FALSE
process = "sst"
sites <- c("Port l'Hebert", "Taylor's Head Shallow")
site_labels <- c("Port L'Hebert", "Taylor's Head")
pal <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442")
source("setup.R")
```

# Reading CMIP6 Data

```{r}
ph_cmip6 <- readCMIP6("../data/Port_LHebert/cmip6_2000_to_2024.mat")
th_cmip6 <- readCMIP6("../data/TaylorHead/cmip6_2000_to_2024.mat")

ph_cmip6 <- ph_cmip6 |>
  mutate("site" = "Port L'Hebert")
th_cmip6 <- th_cmip6 |>
  mutate("site" = "Taylor's Head")
cmip6 <- bind_rows(ph_cmip6, th_cmip6) |>
  mutate("type" = "CMIP6") |>
  select("t" = time, site, type, temp)
```

# Processing Data

Filtering between May 28th, 2018 to March 20th, 2020.

```{r}
hourly <- hourly_dataframe |>
  filter(between(time, dates$freq[1], dates$freq[2])) |>
  select(-GLORYS) |>
  rename("t" = time) |>
  pivot_longer(
    c(OBS, FVCOM, FVCOM2),
    names_to = "type",
    values_to = "temp"
  ) |>
  remove_missing_times()

hourly |>
  ggplot(aes(x = t, y = temp, col = type)) +
  geom_line() +
  facet_wrap(~site, ncol = 1)

daily <- daily_dataframe |>
  filter(between(time, dates$freq[1], dates$freq[2])) |>
  rename("t" = time) |>
  pivot_longer(
    c(OBS, FVCOM, FVCOM2, GLORYS),
    names_to = "type",
    values_to = "temp"
  ) |>
  remove_missing_times()

daily |>
  ggplot(aes(x = t, y = temp, col = type)) +
  geom_line() +
  facet_wrap(~site, ncol = 1)

monthly <- daily |>
  bind_rows(cmip6) |>
  filter(between(t, dates$freq[1], dates$freq[2])) |>
  mutate(
    "year" = year(t),
    "month" = month(t)
  ) |>
  group_by(site, type, year, month) |>
  summarize(
    "temp" = mean(temp),
    .groups = "drop"
  ) |>
  mutate(
    "t" = make_date(year, month),
  ) |>
  select(t, site, type, temp)
  
monthly |>
  ggplot(aes(x = t, y = temp, col = type)) +
  geom_line() +
  facet_wrap(~site, ncol = 1)
```

## Decomposing Temeprature Time Series

```{r}
hourly_filtered <- hourly |>
  group_by(site, type) |>
  mutate(
    "seas" = smooth.spline(temp, nknots = 16)$y,
    "anomaly" = temp - seas,
    "met" = freq_filter(anomaly, 1, 1/48, type = "lowpass"),
    "tidal" = anomaly - met
  ) |>
  ungroup()

daily_filtered <- daily |>
  group_by(site, type) |>
  mutate(
    "seas" = smooth.spline(temp, nknots = 16)$y,
    "anomaly" = temp - seas,
    "met" = freq_filter(anomaly, 1, 1/2, type = "lowpass"),
    "tidal" = anomaly - met
  ) |>
  ungroup()
```

## Combining Hourly, Daily, and Monthly Data

```{r}
all_scales_filtered <- bind_rows(
  hourly_filtered |>
    pivot_longer(
      c(temp, seas, anomaly, met, tidal),
      names_to = "band",
      values_to = "band_temp"
    ) |>
    mutate("scale" = "hourly"),
  daily_filtered |>
    filter(type == "GLORYS") |>
    select(-tidal) |>
    mutate("t" = as.POSIXct(t, tz = "UTC")) |>
    pivot_longer(
      c(temp, seas, anomaly, met),
      names_to = "band",
      values_to = "band_temp"
    ) |>
    mutate("scale" = "daily"),
  monthly |>
    filter(type == "CMIP6") |>
    mutate("t" = as.POSIXct(t, tz = "UTC")) |>
    pivot_longer(
      c(temp),
      names_to = "band",
      values_to = "band_temp"
    ) |>
    mutate("scale" = "monthly")
)
```

# Plotting Raw Temperatures and Seasonal Cycles

```{r}
all_scales_filtered |>
  filter(band %in% c("temp", "seas")) |>
  mutate(type = factor(type, levels = c("OBS", "FVCOM", "FVCOM2", "GLORYS", "CMIP6"))) |>
  mutate(band = factor(band, levels = c("temp", "seas"))) |>
  ggplot(aes(x = t, y = band_temp, col = band, lwd = band)) +
  geom_line() +
  facet_grid(type ~ site, 
             labeller = as_labeller(c("FVCOM" = "FVCOMv1",
                                      "FVCOM2" = "FVCOMv2",
                                      "OBS" = "Observations",
                                      "GLORYS" = "GLORYS12v1",
                                      "CMIP6" = "CMIP6",
                                      "Port L'Hebert" = "Port L'Hebert",
                                      "Taylor's Head" = "Taylor's Head"))) +
  scale_color_manual(
    name = "",
    values = c("black", "blue"),
    breaks = c("temp", "seas"),
    labels = c("Observations", "Seasonal Cycle")
  ) +
  scale_linewidth_manual(
    name = "",
    values = c(0.5, 0.8),
    breaks = c("temp", "seas"),
    labels = c("Observations", "Seasonal Cycle")
  ) +
  #scale_y_continuous(
  #  breaks = c(0, 10, 20)
  #) +
  labs(
    x = "Time",
    y = "Temperature [\u00B0C]"
  ) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )
```

# Plotting Raw Temperature Histograms

```{r}
hourly |>
  mutate(type = factor(type, levels = c("OBS", "FVCOM", "FVCOM2"))) |>
  ggplot(aes(x = temp)) +
  geom_histogram(bins = 40, col = "black", fill = "grey") +
  facet_grid(type ~ site, 
             labeller = as_labeller(c("FVCOM" = "FVCOMv1",
                                      "FVCOM2" = "FVCOMv2",
                                      "OBS" = "Observations",
                                      "Port L'Hebert" = "Port L'Hebert",
                                      "Taylor's Head" = "Taylor's Head"))) +
  theme_bw() +
  labs(
    title = "Histogram of Temperature Data",
    x = "",
    y = "Temperature [\u00B0C]"
  )
```

# Plotting Temperature Anomalies

```{r}
all_scales_filtered |>
  filter(band == "anomaly") |>
  filter(type != "CMIP6") |>
  mutate(type = factor(type, levels = c("OBS", "FVCOM", "FVCOM2", "GLORYS"))) |>
  ggplot(aes(x = t, y = band_temp)) +
  geom_line(col = "black") +
  facet_grid(type ~ site, 
             labeller = as_labeller(c("FVCOM" = "FVCOMv1",
                                      "FVCOM2" = "FVCOMv2",
                                      "OBS" = "Observations",
                                      "GLORYS" = "GLORYS12v1",
                                      "CMIP6" = "CMIP6",
                                      "Port L'Hebert" = "Port L'Hebert",
                                      "Taylor's Head" = "Taylor's Head"))) +
  labs(
    title = "",
    x = "Time",
    y = "Temperature [\u00B0C]"
  ) +
  theme_bw() +
    theme(
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )
```

# Converting to Wide Dataframes

```{r}
ph_hourly_wide <- hourly |>
  filter(site == "Port L'Hebert") |>
  pivot_wider(
    names_from = type,
    values_from = temp
  )

ph_daily_wide <- daily |>
  filter(site == "Port L'Hebert") |>
  pivot_wider(
    names_from = type,
    values_from = temp
  )

ph_monthly_wide <- monthly |>
  filter(site == "Port L'Hebert") |>
  pivot_wider(
    names_from = type,
    values_from = temp
  ) |>
  drop_na()

th_hourly_wide <- hourly |>
  filter(site == "Taylor's Head") |>
  pivot_wider(
    names_from = type,
    values_from = temp
  )

th_daily_wide <- daily |>
  filter(site == "Taylor's Head") |>
  pivot_wider(
    names_from = type,
    values_from = temp
  )

th_monthly_wide <- monthly |>
  filter(site == "Taylor's Head") |>
  pivot_wider(
    names_from = type,
    values_from = temp
  ) |>
  drop_na()
```

# Calculating Traditional Metrics

```{r}
left_join(
  apply_metrics(
    ph_hourly_wide$OBS,
    ph_hourly_wide$FVCOM,
    pred_label = "FVCOMv1",
    metric_funcs = list(bias, aae, mse, rmse, cor, willmott, nse),
    metric_labels = c("B", "AAE", "MSE", "RMSE", "R", "WS", "NSE")
  ),
  apply_metrics(
    ph_hourly_wide$OBS,
    ph_hourly_wide$FVCOM2,
    pred_label = "FVCOMv2",
    metric_funcs = list(bias, aae, mse, rmse, cor, willmott, nse),
    metric_labels = c("B", "AAE", "MSE", "RMSE", "R", "WS", "NSE")
  ),
  by = "metric"
) |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

left_join(
  apply_metrics(
    th_hourly_wide$OBS,
    th_hourly_wide$FVCOM,
    pred_label = "FVCOMv1",
    metric_funcs = list(bias, aae, mse, rmse, cor, willmott, nse),
    metric_labels = c("B", "AAE", "MSE", "RMSE", "R", "WS", "NSE")
  ),
  apply_metrics(
    th_hourly_wide$OBS,
    th_hourly_wide$FVCOM2,
    pred_label = "FVCOMv2",
    metric_funcs = list(bias, aae, mse, rmse, cor, willmott, nse),
    metric_labels = c("B", "AAE", "MSE", "RMSE", "R", "WS", "NSE")
  ),
  by = "metric"
) |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

left_join(
  apply_metrics(
    ph_daily_wide$OBS,
    ph_daily_wide$GLORYS,
    pred_label = "GLORYS12v1",
    metric_funcs = list(bias, aae, mse, rmse, cor, willmott, nse),
    metric_labels = c("B", "AAE", "MSE", "RMSE", "R", "WS", "NSE")
  ),
  apply_metrics(
    ph_monthly_wide$OBS,
    ph_monthly_wide$CMIP6,
    pred_label = "CMIP6",
    metric_funcs = list(bias, aae, mse, rmse, cor, willmott, nse),
    metric_labels = c("B", "AAE", "MSE", "RMSE", "R", "WS", "NSE")
  ),
  by = "metric"
) |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

left_join(
  apply_metrics(
    th_daily_wide$OBS,
    th_daily_wide$GLORYS,
    pred_label = "GLORYS12v1",
    metric_funcs = list(bias, aae, mse, rmse, cor, willmott, nse),
    metric_labels = c("B", "AAE", "MSE", "RMSE", "R", "WS", "NSE")
  ),
  apply_metrics(
    th_monthly_wide$OBS,
    th_monthly_wide$CMIP6,
    pred_label = "CMIP6",
    metric_funcs = list(bias, aae, mse, rmse, cor, willmott, nse),
    metric_labels = c("B", "AAE", "MSE", "RMSE", "R", "WS", "NSE")
  ),
  by = "metric"
) |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))
```

# Plotting Taylor Diagrams

```{r}
taylor_diagram2(
  x = cbind(ph_hourly_wide$OBS, th_hourly_wide$OBS), 
  y = cbind(ph_hourly_wide$FVCOM, ph_hourly_wide$FVCOM2, 
            th_hourly_wide$FVCOM, th_hourly_wide$FVCOM2), 
  sd_max = 1.25, 
  sites = c("Port L'Hebert", "Taylor's Head"), 
  types = c("Observations", "FVCOMv1", "FVCOMv2"), 
  print_results = FALSE
) +
  scale_color_manual(
    name = "",
    values = c("black", pal[1:2]),
    breaks = c("Observations", "FVCOMv1", "FVCOMv2"),
    labels = c("Observations", "FVCOMv1", "FVCOMv2")
  ) +
  scale_x_continuous(
      name = "Normalized standard deviation",
      limits = c(0,1),
      breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1),
      labels = c(" R = 0", " R = 0.2", " R = 0.4",
                 " R = 0.6", " R = 0.8", " R = 1")
  ) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    legend.text = element_text(size = 12)
  )

taylor_diagram2(
  x = cbind(ph_monthly_wide$OBS, th_monthly_wide$OBS), 
  y = cbind(ph_monthly_wide$FVCOM, ph_monthly_wide$FVCOM2,
            ph_monthly_wide$GLORYS, ph_monthly_wide$CMIP6,
            th_monthly_wide$FVCOM, th_monthly_wide$FVCOM2,
            th_monthly_wide$GLORYS, th_monthly_wide$CMIP6),
  sd_max = 1.25, 
  sites = c("Port L'Hebert", "Taylor's Head"), 
  types = c("Observations", "FVCOMv1", "FVCOMv2", "GLORYS12v1", "CMIP6"), 
  print_results = FALSE
) +
  scale_color_manual(
    name = "",
    values = c("black", pal[1:4]),
    breaks = c("Observations", "FVCOMv1", "FVCOMv2", "GLORYS12v1", "CMIP6"),
    labels = c("Observations", "FVCOMv1", "FVCOMv2", "GLORYS12v1", "CMIP6")
  ) +
  scale_shape_manual(
    name = "",
    values = c(16, 17, 15, 18, 7),
    breaks = c("Observations", "FVCOMv1", "FVCOMv2", "GLORYS12v1", "CMIP6"),
    labels = c("Observations", "FVCOMv1", "FVCOMv2", "GLORYS12v1", "CMIP6")
  ) +
  scale_x_continuous(
      name = "Normalized standard deviation",
      limits = c(0,1),
      breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1),
      labels = c(" R = 0", " R = 0.2", " R = 0.4",
                 " R = 0.6", " R = 0.8", " R = 1")
  ) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    legend.text = element_text(size = 12)
  )
```

